---
title: "F1 Racing EDA and Podium Predictions (In Progress)"
author: "Matthew Harris"
date: 11/11/2019
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This will be used as a sample to display some of my analytical capabilities in R. This project will demonstrate how useful R can be to perform analysis that is easy to reproduce and communicate. This is by no means an exhaustive demonstration of my proficiency with R, but should highlight common data analysis functions that I perform regularly.

## Analysis Goals
I have little to no knowledge about F1 racing. My only insight prior to conducting this analysis is that Lewis Hamilton is one of the best drivers to ever touch a steering wheel. My goal for this project is to analyze the data to increase my knowledge of the sport and its participants. After conducting my EDA I plan on creating a model that can be used to predict a driver's probability of finishing 3rd or higher in their next race.

## R Shiny Dashboarding
Coming soon...

## Data Sources
The data used for this analysis contains a wide range of information of information pertaining to F1 races from 1950 through 2017. The data is currently separated into various tables. Part of this exercise will require me to verify the accuracy of the data and combine the tables into a more easy to interpret format. The data used for this analysis can be found at the following Kaggle link. <br>
[F1 Race Data](https://www.kaggle.com/cjgdev/formula-1-race-data-19502017)

## Data Import
Loading necessary packages for analysis.
```{r, message = FALSE}
library(car)
library(tidyverse)
library(lubridate)
library(scales)
library(caret)
library(caretEnsemble)
```

Loading csv files containing the data.
```{r, message = FALSE, warning = FALSE}
circuits <- read_csv("Raw Data/circuits.csv")
constructor_results <- read_csv("Raw Data/constructorResults.csv")
constructors <- read_csv("Raw Data/constructors.csv")
constructor_standings <- read_csv("Raw Data/constructorStandings.csv")
drivers <- read_csv("Raw Data/drivers.csv")
driver_standings <- read_csv("Raw Data/driverStandings.csv")
lap_times <- read_csv("Raw Data/lapTimes.csv")
pit_stops <- read_csv("Raw Data/pitStops.csv")
qualifying <- read_csv("Raw Data/qualifying.csv")
races <- read_csv("Raw Data/races.csv")
results <- read_csv("Raw Data/results.csv")
seasons <- read_csv("Raw Data/seasons.csv")
status <- read_csv("Raw Data/status.csv")
```

## Data Wrangling/Cleansing
### Data Inspection

Now that the data is loaded I can begin to inspect it to determine if any transformations will be needed before I can conduct further analysis. 
```{r}
glimpse(drivers)
```

I will repeat this process with the other tables using the following method.
```{r, eval = FALSE}
map(list(circuits, constructor_results, constructors,
         constructor_standings, drivers, driver_standings,
         lap_times, pit_stops, pit_stops,
         qualifying, races, results,
         seasons, status), glimpse)
```
### Data Cleansing

As expected the data need to be cleaned. There are unneeded fields for links to wikipedia pages, missing data in fields, redundant variables, and datetime issues that need to be resolved. The next couple of sections will focus on the transformations needed to clean the data.

```{r, warning = FALSE}
results %>%
  #Removes the redundant positionText variable
  select(-positionText) %>%
  #Converts the fastestLapTime and race time variables to hms
  mutate(fastestLapTime = hms::as_hms(paste0("00:", str_sub(paste0(fastestLapTime), start = 1, end = 5))),
         f_time = hms::as_hms(as_datetime(milliseconds(milliseconds)))) %>%
  #Removes the updated variables
  select(-time, -milliseconds) -> results

drivers %>%
  #Removing the driverRef variable since it usually the same as the surname
  #Removing the number and code variables becuase they aren't available for all drivers
  select(-c(driverRef, number, code, url)) %>%
  #Converts the dob variable to date format
  mutate(dob = dmy(dob)) -> drivers
```



